---
title: "Final Project"
author: "Dmitriy Ignatov"
date: "`r Sys.Date()`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    dev: pdf
    highlight: tango
  html_document:
    theme: default
    self_contained: true
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    smart: true
    dev: svg
---

```{r setup, include = FALSE}
# DO NOT ALTER THIS CHUNK
# Set knitr options
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.width = 5,
  fig.asp = 0.618,
  out.width = "70%",
  dpi = 120,
  fig.align = "center",
  cache = FALSE
)
# Load required packages
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(infer))
suppressPackageStartupMessages(library(modelr))
suppressPackageStartupMessages(library(broom))
# Load dataset
college <- read_rds("college.rds") %>%
  type_convert(
    na=combine("", "NA", "PrivacySuppressed")
  )
# Set seed
set.seed(98261936)
```

## Introduction

Is there a difference in proportions in degrees awarded after 4 years (graduation) between private and public schools?

* The 2 variables that I will be using are _C100_L4_POOLED_ (continuous) and _SCHTYPE_ (categorical)

* The _C100_L4_POOLED_ variable will be my response variable and the _SCHTYPE_ variable will be my explanatory variable

* An inference hypothesis test will be used in order to see whether a difference in proportions between the two variables exist 

  * The statistical inference and p value from the test statistic will help answer my question

## Preprocessing

```{r}
college_reduced <- college %>%
  select(C100_L4_POOLED,SCHTYPE)
```

* Extracted the two columns I will be using for the project using the _select_ function and saved it into a new variable representing the reduced data set

```{r}
college_reduced <- college_reduced %>%
  rename(completion_rate = C100_L4_POOLED,
         school_type = SCHTYPE)
```

* Used the _rename_ function in order to make the column abbreviations more human readable and clear on what data each column is representing

```{r}
college_reduced <- college_reduced %>%
  mutate(
    school_type = recode(
        school_type,
        `1` = "Public",
        `2` = "Private",
        `3` = "Proprietary"
    )
  )
```

* Used the _mutate_ and _recode_ function in order to give my categorical variable a more appropriate name and relabeled the data whereas the school type is represented by its type rather than a numerical integer

```{r}
college_reduced <- college_reduced %>%
  na.omit(completion_rate)
```

* Used _na.omit_ parameter in order to null any non specified or N/A chunks in data columns 

## Visualization

```{r}
college_reduced %>%
ggplot() +
geom_histogram(
  binwidth = 1,
  position = "identity",
  mapping = aes(x = completion_rate,
  fill = school_type)
  )+
  labs(
    title = "Degrees Awarded by School Type",
    x = "Completion Rate"
  )
```

* Created a histogram visual in order to represent the count of degrees awarded by school types and visualize the co variation

```{r}
college_reduced %>%
  group_by(school_type) %>%
  summarize(
      mean = mean(completion_rate),
      median = median(completion_rate),
      std.dev = sd(completion_rate),
      min     = min(completion_rate),
      max     = max(completion_rate),
      IQR = IQR(completion_rate)
)
```

* Summarized the _completion_rate_ data by school type to further understand it 

```{r}
college_reduced %>%
 ggplot() +
  geom_boxplot(
    mapping = aes(x = completion_rate)) +
  labs(
    title = "Degrees Awarded for Schools"
  )
```

* Created a box plot in order to represent the _completion_rate_ variable and its relation to all schools (not filtered by specific school type)

```{r}
college_reduced %>%
ggplot() +
geom_histogram(
  binwidth = 1,
  position = "identity",
  mapping = aes(x = completion_rate)
  )+
  labs(
    title = "Degrees Awarded by School Type"
    )+
  facet_wrap( ~school_type, scales = "free_x")+
theme(axis.text.x = element_text(angle = 45)) 
```

* Created a histogram visual and faceted on the categorical variable ( _completion_rate_ ) in order to further represent the relationship between both variables

* Creating these visuals allows me to further understand how the data works with one another and helps me answer my question "whether there is a difference in proportions of degrees awarded by school?"

## Summary Statistics

* The test statistic is the difference in the means in the _completion_rate_ variable

* The hypothesis test is two sided as we are looking to test to find the difference in proportions; between two variables

Null Hypothesis: There is no difference in the mean completion rate between public, private, and non proprietary schools

Alternative Hypothesis: There is a difference in mean completion rates between school types (private, public, non proprietary)

```{r}
college_reduced_filtered <- college_reduced %>%
  filter(school_type %in% c("Public", "Private"))
```

```{r}
null_college_reduced <- college_reduced_filtered %>%
  specify(completion_rate ~ school_type) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 10000,  type = "permute") %>%
  calculate(stat = "diff in means",
            order = c("Private", "Public"))
```

* Used a simulation in order to generate the null distribution in order to perform a hypothesis test

```{r}
alt_college_reduced <- college_reduced_filtered %>%
  specify(completion_rate ~ NULL) %>%
  calculate(stat = "mean")
```

* Created an alternative hypothesis test

## Data Analysis

```{r}
stat_college <- college_reduced_filtered %>%
  specify(completion_rate ~ school_type) %>%
  calculate(stat = "diff in means", order = c("Public", "Private"))
```  

* Created a stat variable in order to calculate the p value

```{r}
null_college_reduced %>%
  get_p_value(obs_stat = stat_college, direction = "two_sided")
```

* Calculated the p value of the null hypothesis test

```{r}
null_college_reduced %>%
  visualize() +
  shade_p_value(obs_stat = stat_college, direction = "two_sided"
  )+
labs(
    title = "Null Dsitribution of Degrees Awarded",
    x = "School Type",
    y = "Count"
)                
```

* Created a visual to represent the null distribution simulation which was ran before hand  

(Analyze the hypothesis test and find the test statistic p value if the p value is above .05 it is significantly different between the two variables)


## Conclusion

* Overall the conclusion I am able to draw from this project is that while unfortunately being unable to figure out how to run the hypothesis test in order to see a significance through analyzing the p value. The visuals I created can help me answer the question however not as accurately as wanted.

  * Furthermore, it can be seen that Propreitary schools award more degrees for 4 year full time first students than that of public schools
  
* The analyses from these different sections support each other until the hypotheses testing which I was unable to figure out in the time limit of the project assignment

* I do not believe there are any potentially confounding factors

* This was an interesting question to study because as seen in the visuals, if calculated I believe the p-value as well through the testing. That if someone was selecting a school to go to college/university for and had the goal to pick a school that he has the best percentage chance of being able to graduate in his first 4 years as a first time student, proprietary schools award more degrees than that of public schools. 



