---
title: "Lab 12: Saving for Retirement"
author: "Dmitriy Ignatov"
date: "`r Sys.Date()`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    highlight: tango
    toc: false
    df_print: kable
    fig_caption: no
    number_sections: no
    dev: pdf
  html_document:
    highlight: tango
    self_contained: true
    theme: paper
    toc: no
    df_print: kable
    fig_caption: no
    number_sections: no
    smart: yes
    dev: svg
  github_document:
    html_preview: false
    toc: true
    toc_depth: 4
    dev: svg
    df_print: kable
---

```{r setup, include = FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
# Set knitr options
knitr::opts_chunk$set(
  echo = TRUE, eval = TRUE, fig.width = 5,
  fig.asp = 0.618, out.width = "80%", dpi = 120,
  fig.align = "center", cache = FALSE
)
```

---

## Exercise 1

```{r}
savings <- 0
income <- 58000
interest_rate <- 0.06
savings_rate <- 0.1

for(year in 22:65){
  savings <- savings + 0.06 * savings + 0.1*income 
}
savings
```

* Created a simple simulation in order to track how much we can save for retirement betweeen graduating college and reaching retirement age taking in account income, interest rate, and the savings rate

* Because of the interest rate the savings compound over time and thus the printed amount is so high

## Exercise 2

```{r}
savings * 0.04
```

* Using code calculated how much you would have to live on each year in retirement if you can safely withdraw 4% of the final value of your savings account 

* The annual retirement income is less than our estimated annual salary before retirement

## Exercise 3

```{r}
savings <- 0
income <- 58000
for(year in 22:65){
  savings <- savings + 0.06 * savings + 0.1*income + 0.05*income
}
savings
```
* Using code calculated how much savings would be impacted if the employer matches 5% of your salary

## Exercise 4

```{r}
savings <- 0
income <- 58000 
for(year in 22:65){
  savings <- savings + 0.06 * savings + 0.1*income + 0.05*income
  income <- income + 1000
}
savings
```
* Using code calculated an assumption that our income stays fixed for our whole career and printed out the final savings total

```{r}
savings * 0.04
```
* Using code calculated what 4% interest on this final savings value is 

## Exercise 5

```{r}
savings <- 0
income <- 58000 
for(year in 22:65){
  if(year >= 56){
  savings <- savings + 0.06 * savings + 0.15*income + 0.05*income
}
  income <- income + 1000
}
savings
```
* Using code calculated what would happen if we saved 15% of our income in the last ten years of our career (from ages 56-65, with no savings prior to that)

```{r}
savings <- 0
income <- 58000 
for(year in 22:65){
  if(year <= 31){
  savings <- savings + 0.06 * savings + 0.05*income + 0.05*income
  } else {
savings <- savings + 0.06 * savings 
  }
income <- income + 1000
}
savings
```
* Using code calculated what would happen if we saved 5% of our income in the first 10 years of our career and nothing after that 

* With these results it can be seen it is better to save earlier than later on in life even with a smaller interest/savings rate
